Traffic aware Precoding and Resource Allocation schemes for Multi-Cell \acs{MIMO}-\acs{OFDM} systems

### 2014-09-29 Report/joint_wsrm_formulation.tex
Antti Changes included

The \ac{Q-WSRM} formulation extends the \emph{backpressure algorithm} to the \ac{MIMO}-\ac{OFDM} framework, in which the multiple \acp{BS} acts as the source nodes and the user terminals as the receiver nodes. The control policy in the form of transmit precoders are designed to minimize the number of queued packets waiting at the \acp{BS}. In order to find the optimal algorithm, we use the Lyapunov function which is predominantly used in the control theory for the system stability. Since at each time slot, the system can be described by the channel conditions and the number of backlogged packets of each user, Lyapunov function is used to provide a scalar measure, which grows large when the system moves towards the undesirable state. Following the approach in \cite{neely2010stochastic}, the scalar measure for the queue stability is given by
\begin{equation}
\mathrm{L}\sset{\mbf{Q}(i)} = \frac{1}{2} \, \sum_{k \in \mc{U}} Q_k(i)
\end{equation}
where \me{\mbf{Q}(i)} denotes the stacked user queues at the \me{\ith{i}} slot and \me{\frac{1}{2}} is used for the convenience. The Lyapunov function provides a measure of congestion in the system, as discussed in \cite[Ch. 3]{neely2010stochastic}. Now the Lyapunov function drift is given by
\begin{IEEEeqnarray}{rCl}\label{eqn-3.1}
\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} &=& \frac{1}{2} \left [ \sum_{k \in \mc{U}} \, \Big ( \left [ Q_k(i) - t_k(i) \right ]^+ + \lambda_k(i) \Big )^2 - Q^2_k(i) \right ] \IEEEyessubnumber \label{eqn-3.1.0} \\
&\leq& \sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} + \sum_{k \in \mc{U}} Q_k(i) \set{ \lambda_k(i) - t_k(i) } \IEEEyessubnumber \label{drift-exp}
\end{IEEEeqnarray}
where the inequality is due to the upper bound
\begin{equation}
\sset{\max(Q-t,0) + \lambda)}^2 \leq Q^2 + t^2 + \lambda^2 + 2 Q (\lambda - t)
\end{equation}
In order to minimize the squared sum at each instant, minimization of the Lyapunov drift \eqref{eqn-3.1} is carried over all possible rate allocations in the form of transmission rates \me{t_k} to users in the system. The Lyapunov drift conditioned on the current backlogged packets \me{\mbfa{Q}(i)} is given by
\begin{IEEEeqnarray}{rCl} \label{backpressure-approx}
\underset{\mbf{t}}{\text{minimize}} && \Delta(\mbf{Q}(i)) \triangleq \mathbb{E}_{\mbfa{\lambda},\mbf{t}} \set{\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} \vert \mbf{Q}(i)} \IEEEyessubnumber \\
 &\leq& \mathbb{E}_{\mbfa{\lambda},\mbf{t}} \set {\sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} \vert \mbf{Q}(i)} + \sum_{k \in \mc{U}} Q_k(i) A_k(i) - \mathbb{E}_{\mbfa{\lambda},\mbf{t}}\set{\sum_{k \in \mc{U}} Q_k(i) t_k(i)  \vert \mbf{Q}(i)} \IEEEyessubnumber
\end{IEEEeqnarray}
where the second term on the r.h.s follows from the Poisson arrival process. Assuming the second order moments of the transmission rates and the arrival rates are bounded, it can be replaced by a suitable bound \me{B} in order to eliminate from the optimization problem \cite{neely2010stochastic}.

Now the expression in \eqref{backpressure-approx} looks similar to the \ac{WSRM} formulation if the weights in the \ac{WSRM} problem are replaced by the number of backlogged packets corresponding to the users. The above discussed approach is extended for the wireless networks in \cite{weeraddana2011resource}, where the queue weighted sum rate maximization is considered as the objective function to determine the transmit precoders. The \ac{Q-WSRM} formulation is given by
\begin{IEEEeqnarray}{rCl} \label{q_gen_sum}
	\underset{\substack{\mvec{M}{k,n}, \mvec{W}{k,n}}}{\text{maximize}} &\quad& \sum_{k \in \mc{U}} \, Q_k \left ( \sum_{n = 1}^N \sum_{l=1}^L \log_2(1 + \gamma_{l,k,n}) \right ) \IEEEyessubnumber \label{eqn-3.1.1} \\
	\text{subject to.} & \quad & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \text{tr} \, (\mvec{M}{k,n} \mvec{M}{k,n}^\herm) \leq P_{{\max}}, \fall b \IEEEyessubnumber \label{eqn-3.1.3}
\end{IEEEeqnarray}
In order to avoid the excessive  allocation of the resources, we include an additional rate constraint \me{t_k \leq Q_k} to address \me{[x]^+} operation in \eqref{eqn-2a}. The rate constrained version of the \ac{Q-WSRM}, denoted by \ac{Q-WSRME} problem for a cellular system, is given by \label{q_gen_sum-1} with the additional constraint
\begin{IEEEeqnarray}{rCl} \label{eqn-3.1.4}
\sum_{n=1}^N \sum_{l = 1}^L \log_2(1 + \gamma_{l,k,n}) \leq Q_k, \fall k \in \mc{U}
\end{IEEEeqnarray}
where the precoders are associated with the \me{\gamma_{l,k,n}} defined in \eqref{eq:SINR}. By using the number of queued packets as the weights, the resources can be allocated to the user with the more number of backlogged packets, which essentially does the allocation in a greedy manner.

As a special case of the problem defined in \eqref{q_gen_sum}, we can formulate the sum rate maximization problem by setting the weights in \eqref{eqn-3.1.1} as unity, leading to the problem as
\begin{IEEEeqnarray}{rCl}\label{gen_sum}
\underset{\substack{\mvec{M}{k,n},\mvec{W}{k,n}}}{\text{maximize}} &\quad& \sum_{k \in \mc{U}} \, \sum_{n = 1}^N \sum_{l=1}^L \log_2(1 + \gamma_{l,k,n}) \IEEEyessubnumber \\
\text{subject to.} & \quad & \eqref{eqn-3.1.3} \; \text{and} \; \eqref{eqn-3.1.4} \IEEEyessubnumber
\end{IEEEeqnarray}
The problem defined in \eqref{gen_sum} provides a greedy queue minimizing approach as compared to \ac{Q-WSRME}, since the resource allocation is driven by the channel conditions in comparison with the number of queued packets as in \ac{Q-WSRME}. Note that in both formulations, the resources allocated to the users are limited by the backlogged packets with an explicit maximum rate constraint defined by \eqref{eqn-3.1.4}.

>>>>

The \ac{Q-WSRM} formulation extends the \emph{backpressure algorithm} to the \ac{MIMO}-\ac{OFDM} framework, in which the multiple \acp{BS} acts as the source nodes and the user terminals as the receiver nodes. The control policy in the form of transmit precoders are designed to minimize the number of queued packets waiting at the \acp{BS}. In order to find the optimal algorithm, we use the Lyapunov function which is predominantly used in the control theory for the system stability. Since at each time slot, the system can be described by the channel conditions and the number of backlogged packets of each user, Lyapunov function is used to provide a scalar measure, which grows large when the system moves towards the undesirable state. Following the approach in \cite{neely2010stochastic}, the scalar measure for the queue stability is given by
\begin{equation}
\mathrm{L}\sset{\mbf{Q}(i)} = \frac{1}{2} \, \sum_{k \in \mc{U}} Q_k^2(i)
\end{equation}
where \me{\mbf{Q}(i)} denotes the stacked user queues at the \me{\ith{i}} slot and \me{\frac{1}{2}} is used for the convenience. The Lyapunov function provides a measure of congestion in the system, as discussed in \cite[Ch. 3]{neely2010stochastic}. Now the Lyapunov function drift is given by
\begin{IEEEeqnarray}{rCl}\label{eqn-3.1}
\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} &=& \frac{1}{2} \left [ \sum_{k \in \mc{U}} \, \Big ( \left [ Q_k(i) - t_k(i) \right ]^+ + \lambda_k(i) \Big )^2 - Q^2_k(i) \right ] \IEEEyessubnumber \label{eqn-3.1.0} \\
&\leq& \sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} + \sum_{k \in \mc{U}} Q_k(i) \set{ \lambda_k(i) - t_k(i) } \IEEEyessubnumber \label{drift-exp}
\end{IEEEeqnarray}
where the inequality is due to the upper bound
\begin{equation}
\sset{\max(Q-t,0) + \lambda)}^2 \leq Q^2 + t^2 + \lambda^2 + 2 Q (\lambda - t)
\end{equation}
In order to minimize the squared sum at each instant, minimization of the Lyapunov drift \eqref{eqn-3.1} is carried over all possible rate allocations in the form of transmission rates \me{t_k} to users in the system. The Lyapunov drift conditioned on the current backlogged packets \me{\mbfa{Q}(i)} is given by
\begin{IEEEeqnarray}{rCl} \label{backpressure-approx}
\underset{\mbf{t}}{\text{minimize}} && \Delta(\mbf{Q}(i)) \triangleq \mathbb{E}_{\mbfa{\lambda},\mbf{t}} \set{\mathrm{L}\sset{\mbf{Q}(i+1)} - \mathrm{L}\sset{\mbf{Q}(i)} \vert \mbf{Q}(i)} \IEEEyessubnumber \\
 &\leq& \underbrace{\mathbb{E}_{\mbfa{\lambda},\mbf{t}} \set {\sum_{k \in \mc{U}} \, \frac{\lambda^2_k(i) + t_k^2(i)}{2} \vert \mbf{Q}(i)}}_{\le B} + \sum_{k \in \mc{U}} Q_k(i) A_k(i) - \mathbb{E}_{\mbfa{\lambda},\mbf{t}}\set{\sum_{k \in \mc{U}} Q_k(i) t_k(i)  \vert \mbf{Q}(i)} \IEEEyessubnumber \label{backpreassure-approx-1b}
\end{IEEEeqnarray}
where the second term on the r.h.s follows from the Poisson arrival process. Assuming the second order moments of the transmission rates and the arrival rates are bounded, the sum of arrivals and the transmission rates can be replaced by a upper bound \me{B} as in \eqref{backpreassure-approx-1b}, in order to eliminate from the optimization problem \cite{neely2010stochastic}. The expectation is performed across all possible arrivals and the transmission rates for the given number of queued packets \me{\mbf{Q}(i)} and the channel state information at the \me{\ith{i}} slot.

Now the expression in \eqref{backpressure-approx} looks similar to the \ac{WSRM} formulation if the weights in the \ac{WSRM} problem are replaced by the number of backlogged packets corresponding to the users. The above discussed approach is extended for the wireless networks in \cite{weeraddana2011resource}, where the queue weighted sum rate maximization is considered as the objective function to determine the transmit precoders. The \ac{Q-WSRM} formulation is given by optimizing each term inside the expectation operator \me{\mbf{E}_{\mbfa{\lambda},\mbf{t}}} in \eqref{backpreassure-approx-1b}, which maximizes the 
\begin{IEEEeqnarray}{rCl} \label{q_gen_sum}
	\underset{\substack{\mvec{M}{k,n}, \mvec{W}{k,n}}}{\text{maximize}} &\quad& \sum_{k \in \mc{U}} \, Q_k \left ( \sum_{n = 1}^N \sum_{l=1}^L \log_2(1 + \gamma_{l,k,n}) \right ) \IEEEyessubnumber \label{eqn-3.1.1} \\
	\text{subject to.} & \quad & \sum_{n = 1}^N \sum_{k \in \mathcal{U}_b} \text{tr} \, (\mvec{M}{k,n} \mvec{M}{k,n}^\herm) \leq P_{{\max}}, \fall b \IEEEyessubnumber \label{eqn-3.1.3}
\end{IEEEeqnarray}
In order to avoid the excessive  allocation of the resources, we include an additional rate constraint \me{t_k \leq Q_k} to address \me{[x]^+} operation in \eqref{eqn-2a}. The rate constrained version of the \ac{Q-WSRM}, denoted by \ac{Q-WSRME} problem for a cellular system, is given by \label{q_gen_sum-1} with the additional constraint
\begin{IEEEeqnarray}{rCl} \label{eqn-3.1.4}
\sum_{n=1}^N \sum_{l = 1}^L \log_2(1 + \gamma_{l,k,n}) \leq Q_k, \fall k \in \mc{U}
\end{IEEEeqnarray}
where the precoders are associated with the \me{\gamma_{l,k,n}} defined in \eqref{eq:SINR}. By using the number of queued packets as the weights, the resources can be allocated to the user with the more number of backlogged packets, which essentially does the allocation in a greedy manner.

As a special case of the problem defined in \eqref{q_gen_sum}, we can formulate the sum rate maximization problem by setting the weights in \eqref{eqn-3.1.1} as unity, leading to the problem as in \eqref{q_gen_sum} with \me{Q_k = 1, \forall k \in \mc{U}}. This approach provides a greedy queue minimizing allocation as compared to \ac{Q-WSRME}, since the resource allocation is driven by the channel conditions in comparison with the number of queued packets as in \ac{Q-WSRME}. Note that in both formulations, the resources allocated to the users are limited by the backlogged packets with an explicit maximum rate constraint defined by \eqref{eqn-3.1.4}.

### 2014-09-29 Report/kkt_mse_approach.tex
Antti Changes included

The distributed solutions via primal and \ac{ADMM} approaches depend on the subgradient update by using a step size parameter for the coupling variables, which affects the speed of convergence to the optimal value. In this method, we provide an alternative approach to decentralize the \ac{MSE} equivalent problem considered in \cite{christensen2008weighted,wmmse_shi} by solving the \ac{KKT} conditions. Similar work has been considered for the \ac{WSRM} problem with the minimum rate constraints in \cite{kaleva2013decentralized,kaleva2013primal}. When the queues are involved, the maximum rate constraint imposed by the number of queued packets at the \ac{BS} includes a nonconvex constraint, which makes the problem difficult to solve due to the additional nonconvex maximum rate constraint \eqref{eqn-3.1.4} for the \ac{WSRM} problem.

@@

By forming the Lagrangian of \eqref{kkt-mse-1} with the corresponding dual variables as shown in \eqref{kkt-mse-1}, we can obtain the \ac{KKT} expressions by differentiating with respect to the variables present in the problem as detailed in Appendix \ref{a-1}. Upon solving the \ac{KKT} expressions, the iterative solution \me{\forall k \in \mc{U}}, \me{\forall n \in \set{1,\ldots,N}} and \me{\forall l \in \set{1,\ldots,L}} is given by
\begin{IEEEeqnarray}{rCl} \label{kkt-mse-4}
\mvec{m}{l,k,n}^{(i)} &=& \Big ( \sum_{x \in \mc{U}} \sum_{y=1}^L \alpha_{y,x,n}^{(i-1)} \mvec{H}{b_k,x,n}^\herm \mvec{w}{y,x,n}^{(i-1)} \mvec{w}{y,x,n}^{\herm \, {(i-1)}} \mvec{H}{b_k,x,n} + \delta_b \mbf{I}_{N_T} \Big )^{-1} \alpha^{(i-1)}_{l,k,n} \mvec{H}{b_k,k,n}^\herm \mvec{w}{l,k,n}^{(i-1)} \IEEEyessubnumber \label{kkt-mse-4.3} \\
\epsilon_{l,k,n}^{(i)} &=& \left | 1 - \mvec{w}{l,k,n}^{\herm \, (i)} \mvec{H}{b_k,k,n} \mvec{m}{l,k,n}^{(i)} \right |^2 + \sum_{\mathclap{(x,y) \neq (l,k)}} \left | \mvec{w}{l,k,n}^{\mathrm{H} \, (i)} \mvec{H}{b_y,y,n} \mvec{m}{x,y,n}^{(i)} \right |^2 + N_0 \, \|\mvec{w}{l,k,n}^{(i)}\|^2 \IEEEyessubnumber \label{kkt-mse-4.4} \\
t_{l,k,n}^{(i)} &=&  -\log_2(\epsilon_{l,k,n}^{(i-1)}) - \frac{\left ( \epsilon_{l,k,n}^{(i)} - \epsilon_{l,k,n}^{(i-1)} \right ) }{\log(2) \, \epsilon_{l,k,n}^{(i-1)}} \IEEEyessubnumber \label{kkt-mse-4.5} \\
\sigma_{l,k,n}^{(i)} &=& \max{\left \lbrace 0, \frac{a_k \, q}{\log(2)}  \, \Big (Q_k - \sum_{n = 1}^N \sum_{l=1}^L t_{l,k,n}^{(i)} \Big )^{(q-1)}\right \rbrace }  \IEEEyessubnumber \label{kkt-mse-4.2} \\
\alpha^{(i)}_{l,k,n} &=& \alpha^{(i-1)}_{l,k,n} + \rho \left ( \dfrac{\sigma_{l,k,n}^{(i)}}{\epsilon_{l,k,n}^{(i)}} - \alpha^{(i-1)}_{l,k,n} \right ) \IEEEyessubnumber \label{kkt-mse-4.1} \\
\mvec{w}{l,k,n}^{(i)} &=& \Big ( \sum_{x\in\mc{U}}\sum_{y=1}^L \mvec{H}{b_x,k,n} \mvec{m}{y,x,n}^{(i)} \mvec{m}{y,x,n}^{\herm \, (i)} \mvec{H}{b_{x},k,n}^\herm + \mathbf{I}_{N_R} \Big ) ^{-1} \; \mvec{H}{b_k,k,n} \; \mvec{m}{l,k,n}^{(i)} \IEEEyessubnumber \label{kkt-mse-4.6}
\end{IEEEeqnarray}
Since the dual variables \me{\alpha^{(i)}} and \me{\sigma^{(i)}} are interdependent in \eqref{kkt-mse-4}, one has to be fixed to optimize for the other. In this problem, we fix the dual variable \me{\alpha^{(i)}} to a fixed value as in \eqref{kkt-mse-4.1} to obtain the other variables in \eqref{kkt-mse-4}. It can be seen that, when the allocated rate \me{t_k^{(i-1)}} is greater than the number of queued packets \me{Q_k} for a user \me{k}, the corresponding dual variable \me{\sigma^{(i)}} will be negative thereby forcing the dual variable \me{\alpha_k^{(i)} < \alpha_k^{(i-1)}}, which in turn reduce the transmit precoder weights in \eqref{kkt-mse-4.3}. This reduction in the precoder weights forces the allocated rate of the user \me{k} to reduce from the previous iteration as \me{t_k^{(i)} < t_k^{(i-1)}}.

@@

\subsubsection*{Convergence}
The iterative method presented in Algorithm \ref{algo-4} converges to the stationary point if the dual variables \me{\alpha_{l,k,n}} are allowed to converge. The convergence of the dual variable is guaranteed, since the problem is convex by fixing the receive precoders \me{\mvec{w}{l,k,n}} and the operating \ac{MSE} point \me{\epsilon_{l,k,n}} \cite{boyd2011distributed}. Once the dual variables are converged or iterated to a certain accuracy, the receivers are updated using the \ac{MMSE} objective. In this algorithm, when \me{t_k > Q_k}, \me{\sigma_{l,k,n}} will be zero (dual feasibility), thereby reducing the priority weights \me{\alpha_{l,k,n}} present in the transmit precoder expression in \eqref{kkt-mse-4.3}.

>>>>

The distributed solutions via primal and \ac{ADMM} approaches depend on the subgradient update by using a step size parameter for the coupling variables, which affects the speed of convergence to the optimal value. In this method, we provide an alternative approach to decentralize the \ac{MSE} equivalent problem considered in \cite{christensen2008weighted,wmmse_shi} by directly solving the \ac{KKT} conditions. Similar work has been considered for the \ac{WSRM} problem with the minimum rate constraints in \cite{kaleva2013decentralized,kaleva2013primal}. When the queues are involved, the maximum rate constraint imposed by the number of queued packets at the \ac{BS} includes a nonconvex constraint, which makes the problem difficult to solve due to the additional nonconvex maximum rate constraint \eqref{eqn-3.1.4} for the \ac{WSRM} problem.

@@

By forming the Lagrangian of \eqref{kkt-mse-1} with the corresponding dual variables as shown in \eqref{kkt-mse-1}, we can obtain the \ac{KKT} expressions by differentiating with respect to the variables present in the problem as detailed in Appendix \ref{a-1}. Upon solving the \ac{KKT} expressions, the iterative solution \me{\forall k \in \mc{U}}, \me{\forall n \in \set{1,\ldots,N}} and \me{\forall l \in \set{1,\ldots,L}} is given by
\begin{IEEEeqnarray}{rCl} \label{kkt-mse-4}
\mvec{m}{l,k,n}^{(i)} &=& \Big ( \sum_{x \in \mc{U}} \sum_{y=1}^L \alpha_{y,x,n}^{(i-1)} \mvec{H}{b_k,x,n}^\herm \mvec{w}{y,x,n}^{(i-1)} \mvec{w}{y,x,n}^{\herm \, {(i-1)}} \mvec{H}{b_k,x,n} + \delta_b \mbf{I}_{N_T} \Big )^{-1} \alpha^{(i-1)}_{l,k,n} \mvec{H}{b_k,k,n}^\herm \mvec{w}{l,k,n}^{(i-1)} \IEEEyessubnumber \label{kkt-mse-4.3} \\
\epsilon_{l,k,n}^{(i)} &=& \left | 1 - \mvec{w}{l,k,n}^{\herm \, (i)} \mvec{H}{b_k,k,n} \mvec{m}{l,k,n}^{(i)} \right |^2 + \sum_{\mathclap{(x,y) \neq (l,k)}} \left | \mvec{w}{l,k,n}^{\mathrm{H} \, (i)} \mvec{H}{b_y,y,n} \mvec{m}{x,y,n}^{(i)} \right |^2 + N_0 \, \|\mvec{w}{l,k,n}^{(i)}\|^2 \IEEEyessubnumber \label{kkt-mse-4.4} \\
t_{l,k,n}^{(i)} &=&  -\log_2(\epsilon_{l,k,n}^{(i-1)}) - \frac{\left ( \epsilon_{l,k,n}^{(i)} - \epsilon_{l,k,n}^{(i-1)} \right ) }{\log(2) \, \epsilon_{l,k,n}^{(i-1)}} \IEEEyessubnumber \label{kkt-mse-4.5} \\
\sigma_{l,k,n}^{(i)} &=& \left [\frac{a_k \, q}{\log(2)}  \, \Big (Q_k - \sum_{n = 1}^N \sum_{l=1}^L t_{l,k,n}^{(i)} \Big )^{(q-1)}\right ]^+  \IEEEyessubnumber \label{kkt-mse-4.2} \\
\alpha^{(i)}_{l,k,n} &=& \alpha^{(i-1)}_{l,k,n} + \rho \left ( \dfrac{\sigma_{l,k,n}^{(i)}}{\epsilon_{l,k,n}^{(i)}} - \alpha^{(i-1)}_{l,k,n} \right ) \IEEEyessubnumber \label{kkt-mse-4.1} \\
\mvec{w}{l,k,n}^{(i)} &=& \Big ( \sum_{x\in\mc{U}}\sum_{y=1}^L \mvec{H}{b_x,k,n} \mvec{m}{y,x,n}^{(i)} \mvec{m}{y,x,n}^{\herm \, (i)} \mvec{H}{b_{x},k,n}^\herm + \mathbf{I}_{N_R} \Big ) ^{-1} \; \mvec{H}{b_k,k,n} \; \mvec{m}{l,k,n}^{(i)} \IEEEyessubnumber \label{kkt-mse-4.6}
\end{IEEEeqnarray}
Since the dual variables \me{\alpha^{(i)}} and \me{\sigma^{(i)}} are interdependent in \eqref{kkt-mse-4}, one has to be fixed to optimize for the other. In this problem, we fix the dual variable \me{\alpha^{(i)}} to a fixed value as in \eqref{kkt-mse-4.1} to obtain the other variables in \eqref{kkt-mse-4}. At each iteration, the dual variable \me{\alpha^{(i)}} is updated linearly from the earlier value \me{\alpha^{(i-1)}} by a step size of \me{\rho \in [0,1]}. It can be seen that, when the allocated rate \me{t_k^{(i-1)}} is greater than the number of queued packets \me{Q_k} for a user \me{k}, the corresponding dual variable \me{\sigma^{(i)}} will be negative thereby forcing the dual variable \me{\alpha_k^{(i)} < \alpha_k^{(i-1)}}, which in turn reduce the transmit precoder weights in \eqref{kkt-mse-4.3}. This reduction in the precoder weights forces the allocated rate of the user \me{k} to reduce from the previous iteration as \me{t_k^{(i)} < t_k^{(i-1)}}.

@@

\subsubsection*{Convergence}
The iterative method presented in Algorithm \ref{algo-4} converges to the stationary point if the dual variables \me{\alpha_{l,k,n}} are allowed to converge or for fixed number of iterations \me{J_{\max}}. The convergence of the dual variable is guaranteed, since the problem is convex by fixing the receive precoders \me{\mvec{w}{l,k,n}} and the operating \ac{MSE} point \me{\epsilon_{l,k,n}} \cite{boyd2011distributed}. Once the dual variables are converged or iterated to a certain accuracy, the receivers are updated using the \ac{MMSE} objective. In this algorithm, when \me{t_k > Q_k}, \me{\sigma_{l,k,n}} will be zero (dual feasibility), thereby reducing the priority weights \me{\alpha_{l,k,n}} present in the transmit precoder expression in \eqref{kkt-mse-4.3}.

### 2014-09-29 Report/mse_formulation.tex
Antti Changes included

\subsubsection*{Convergence}
Following the similar arguments as in  Section \ref{sec-3.2.1}, we can see that the \ac{SCA} subproblems converges to a stationary point. The transmit precoders obtained by the \ac{MSE} relaxations are unique due to the \ac{MSE} expression in \eqref{eqn-mse-2.3}, since it removes the phase ambiguity due to the real valued expression for the transmit precoders as
\begin{equation}
\epsilon_{l,k,n} \geq  1 - 2 \Re{ \left \lbrace \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right \rbrace} + \sum_{\mathclap{\forall (j,i)}} \left | \mvec{w}{l,k,n}^\herm \mvec{H}{b_i,i,n} \mvec{m}{j,i,n} \right |^2 + N_0 \, \|\mvec{w}{l,k,n}\|^2
\end{equation}
At each \ac{SCA} step, the transmit precoders are obtained uniquely by minimizing the weighted \ac{MSE} criterion due to the convex nature of the relaxed problem. Upon finding the optimal transmit precoders, the receive beamformers are updated using the \ac{MMSE} receivers. The \ac{MMSE} receiver minimizes the objective value for the fixed transmit precoders thereby leading to the monotonic convergence of the objective function. At each \ac{SCA} step, the optimal value of the previous iteration is also included in the domain of the problem, the objective value can either decrease or stays the same after each iteration.

>>>>

\subsubsection*{Convergence}
Following the similar arguments as in  Section \ref{sec-3.2.1}, we can see that the \ac{SCA} subproblems converges to a stationary point of the original nonconvex problem at each iteration. The uniqueness of the transmit and the receive precoders are justified if the phase ambiguity from the stationary solution is eliminated. It can be seen from the \ac{MSE} expression \eqref{eqn-mse-2.3} that the phase rotations are fixed for both transmit and the receive precoders at the stationary point, which is evident from the following reformulation of \eqref{eqn-mse-2.3} as
\begin{equation}
\epsilon_{l,k,n} \geq  1 - 2 \Re{ \left \lbrace \mvec{w}{l,k,n}^\herm \mvec{H}{b_k,k,n} \mvec{m}{l,k,n} \right \rbrace} + \sum_{\mathclap{\forall (j,i)}} \left | \mvec{w}{l,k,n}^\herm \mvec{H}{b_i,i,n} \mvec{m}{j,i,n} \right |^2 + N_0 \, \|\mvec{w}{l,k,n}\|^2
\end{equation}
At each \ac{SCA} step, the transmit precoders are obtained uniquely by minimizing the weighted \ac{MSE} criterion due to the convex nature of the relaxed problem. Upon finding the optimal transmit precoders, the receive beamformers are updated using the \ac{MMSE} receivers. The \ac{MMSE} receiver minimizes the objective value for the fixed transmit precoders thereby leading to the monotonic convergence of the objective function. At each \ac{SCA} step, the optimal value of the previous iteration is also included in the domain of the problem, the objective value can either decrease or stays the same after each iteration.

### 2014-09-29 Report/primal_approach.tex
Antti Changes included

The \acl{PD} approach decomposes the problem by fixing the interference variables \me{\zeta_{l,k,n,b} \forall k,b} in order to perform the precoder design independently across each \ac{BS}. Once the optimal precoders are designed at each \ac{BS} with the fixed interference constraints \eqref{eqn-decent-3a}, the dual variables corresponding to the interference constraints are exchanged between the cooperating \acp{BS} \me{\mc{B}} to update the interference variables \me{\zeta_{l,k,n,b}} for the next iteration and is carried out until convergence. The primal approach is discussed extensively for the min-power problem in \cite{pennanen2011decentralized} and much of the current work follows the same. Much of the details are provided in the Appendix \ref{a-2}.

\subsubsection*{Convergence}
The convergence of the primal is similar to that of the centralized problem if the interference variables \me{\zeta_{l,k,n,b}} are allowed to converge to the optimal values. But in practice, we can limit the number of exchanges to \me{J_{\max}} after which \ac{SCA} update is performed until convergence or for \me{I_{\max}} times. The update of \me{\tilde{p}_{l,k,n}, \tilde{q}_{l,k,n}} and \me{\tilde{\beta}_{l,k,n}} can be made in conjunction with the receiver update \me{\mvec{W}{k,n}}. The receiver update can be made by using the precoded pilot transmission from each user as in \cite{komulainen2013effective}.

>>>>

The \acl{PD} approach decomposes the problem by fixing the interference variables \me{\zeta_{l,k,n,b} \forall k,b} in order to perform the precoder design independently across each \ac{BS}. Once the optimal precoders are designed at each \ac{BS} with the fixed interference constraints \eqref{eqn-decent-3a}, the dual variables corresponding to the interference constraints are exchanged between the cooperating \acp{BS} \me{\mc{B}} to update the interference variables \me{\zeta_{l,k,n,b}} for the next iteration and is carried out until convergence. The primal approach is discussed extensively for the min-power problem in \cite{pennanen2011decentralized} and much of the current work follows the same. 
%Much of the details are provided in the Appendix \ref{a-2}.

\subsubsection*{Convergence}
The convergence of the primal is similar to that of the centralized problem if the interference variables \me{\zeta_{l,k,n,b}} are allowed to converge to a stationary point. But in practice, we can limit the number of exchanges to \me{J_{\max}} after which \ac{SCA} update is performed until convergence or for \me{I_{\max}} times. The update of \me{\tilde{p}_{l,k,n}, \tilde{q}_{l,k,n}} and \me{\tilde{\beta}_{l,k,n}} can be made in conjunction with the receiver update \me{\mvec{W}{k,n}}. The receiver update can be made by using the precoded pilot transmission from each user as in \cite{komulainen2013effective}.

### 2014-09-29 Report/problem_formulation.tex
Antti Changes included

The problem defined in \eqref{q_gen_sum} ignores the second order term arising from the Lyapunov drift minimization objective by the limiting it to a constant value. In fact, Eq. \eqref{eqn-4.2} provides similar expression when the exponent is set to be \me{\ell_{q=2}} as
\begin{equation}\label{pf-1}
\underset{t_k}{\text{minimize}} \, \sum_k \, v_k^2 = \underset{t_k}{\text{minimize}} \, \sum_k \, Q_k^2 - 2 \, Q_k t_k + t_k^2
\end{equation}
It is evident that \eqref{pf-1} is equivalent to \eqref{backpressure-approx} if the second order terms are ignored. Limiting \me{t_k^2} by a constant value, the \ac{Q-WSRM} formulation requires the explicit rate constraint \eqref{eqn-3.1.4} to avoid the resource wastage in the form of over allocation. In the current queue deviation formulation, the explicit rate constraint is not needed, since it is handled by the objective function itself. In contrast to the \ac{WSRM} formulation, the \ac{JSFRA} and the \ac{Q-WSRM} problems include the sub-channels jointly to achieve an efficient allocation by identifying the optimal space-frequency resource for each user in the system. The queue deviation objective provides an alternative approach to perform the resource allocation without the additional rate constraints as in \ac{Q-WSRME} approach.

>>>>

The problem defined in \eqref{q_gen_sum} ignores the second order term arising from the Lyapunov drift minimization objective by the limiting it to a constant value. In fact, Eq. \eqref{eqn-4.2} provides similar expression when the exponent is set to be \me{\ell_{q=2}} as
\begin{equation}\label{pf-1}
\underset{t_k}{\text{minimize}} \, \sum_k \, v_k^2 = \underset{t_k}{\text{minimize}} \, \sum_k \, Q_k^2 - 2 \, Q_k t_k + t_k^2
\end{equation}
It is evident that \eqref{pf-1} is equivalent to \eqref{backpressure-approx} if the second order terms are ignored. Limiting \me{t_k^2} by a constant value, the \ac{Q-WSRM} formulation requires the explicit rate constraint \eqref{eqn-3.1.4} to avoid the resource wastage in the form of over allocation. In the proposed queue deviation formulation, the explicit rate constraint is not needed, since it is handled by the objective function itself. This makes the problem simpler and allows us to employ efficient algorithms to distribute the precoder design problem across each \acp{BS} independently by exchanging minimal information exchange \cite{boyd2011distributed}. In contrast to the \ac{WSRM} formulation, the \ac{JSFRA} and the \ac{Q-WSRM} problems include the sub-channels jointly to achieve an efficient allocation by identifying the optimal space-frequency resource for each user in the system. The queue deviation objective provides an alternative approach to perform the resource allocation without the additional rate constraints as in \ac{Q-WSRME} approach.
